---
title: "Statistical assignment 1"
author: "[Alhoussein Nasser Candidate number: 142905]"
date: "[add date here]"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Open data (10 points)

In this assignment you will work with the individual level data from wave 8 of the Understanding Society survey. First, you need to open the data set. Please complete the code below.

```{r}
library(tidyverse)
Data <- read_tsv("/Users/alhousseinnasser/Downloads/UKDA-6614-tab 2/tab/ukhls_w8/h_indresp.tab")
view(Data)
```

Now you have got your data frame stored as Data. 

## Select variables (10 points)

The data for Wave 8 of the Understanding Society were collected in 2016-18. Among other things, people were asked the following question: "Should the United Kingdom remain a member of the European Union or leave the European Union?" In this assignment, we will explore how answers to this question depend on sex and age.

First, you need to select the variables for the analysis. You want to keep the following variables: cross-wave individual identifier (*pidp*), support for the UK remaining or leaving the EU (*h_eumem*), sex (*h_sex_dv*), age (*h_age_dv*), and sample origin (*h_memorig*).

Complete the code below to select those variables from the data frame and save the result.

```{r}
Data <- Data %>%
        select(pidp, h_eumem, h_sex_dv, h_age_dv, h_memorig)    
```

## Filter observations (10 points)

To make nationally representative estimates from the Understanding Society data we would need to use weight coefficients. There are many different types of weight coefficients that can be used depending on the question and the level of analysis (see the User Guide, pp. 65-71). We will not do this in this assignment. However, what we want to do is to keep data from the original Understanding Society sample only (ukhls gb 2009-10), dropping  data for Northern Ireland, the BHPS cohort members and ethnic minority boost samples. This will make data closer to be representative for Great Britain. You need to choose the observations where *h_memorig* has the value of 1.

```{r}
Data <- Data %>%
        filter(h_memorig == 1)
```

## Recode data (20 points)

Let us tabulate the variables for EU support, sex, and age.

```{r}
table(Data$h_eumem)
table(Data$h_sex_dv)
table(Data$h_age_dv)
```

You will see that all these variables are numeric. You can learn what the numeric codes mean by checking the codebook here: https://www.understandingsociety.ac.uk/documentation/mainstage/dataset-documentation/datafile/indresp/wave/8 .

We want to do the following:

1) Recode the variable for EU support as binary (1 for Remain, 0 for Leave), coding all types of missing values (including refusals and "don't know") as NA.
2) Recode sex into a character vector with the values "male" or "female".
3) Recode age into a variable with the following categories: 16 to 25, 26 to 40, 41 to 55, 56 to 70, over 70.

In all cases, we want to create new variables.

```{r}
Data <- Data %>%
        mutate(EU = case_when( 
          h_eumem == 1 ~ 1,
          h_eumem == 2 ~ 0)
 
      
        ) %>%
        mutate(sex = case_when(
          h_sex_dv == 1 ~ "male", 
          h_sex_dv == 2 ~ "female"
        )) %>%
        mutate(agegr = case_when(
              between(h_age_dv, 16, 25) ~ "16 to 25",
              between(h_age_dv, 26, 40) ~ "26 to 40",
              between(h_age_dv, 41, 55) ~ "41 to 55",
              between(h_age_dv, 56, 70) ~ "56 to 70",
              h_age_dv > 70 ~ "over 70"
        )  
             
        )

```

## Summarise data (20 points)

Let us **dplyr** to calculate how many people in the sample supported Remain and Leave, both as absolute numbers and percentages.

```{r}
Data %>%  filter(EU > -1 )%>%
     count(EU) %>%
  mutate(perc = (n / sum(n)) * 100)   
        
```

Write a couple of sentences with the interpretation of this result. How this compares with the result of the 2016 referendum? Why?    
45.6492% voted to leave, 54.3508% voted to remain. Assuming that the unavailable data I ignored would have not affected the overall percentages by a significant margin, the results show a majority vote to remain in the EU with a difference of over 8%. This is a more decisive difference when compared to the 2016 referendum which resulted in a majority vote for leaving the au by a margin of over 3%. This shows a notable shift in opinion since 2016.

## Summarise data by sex and age (30 points)

Now let us look at the support for Leave and Remain by sex and age. Use your newly created variables.


```{r}
Data %>%  
  filter(EU >-1 & (sex == "male" | sex == "female")) %>%
  count( sex, EU, agegr) %>% 
  mutate(perc = (n / sum(n)) * 100)

Data %>%  
  filter(EU >-1 & (sex == "female")) %>%
  count( sex, EU) %>% 
  mutate(perc = (n / sum(n)) * 100)

Data %>%  
  filter(EU >-1 & (sex == "male")) %>%
  count( sex, EU) %>% 
  mutate(perc = (n / sum(n)) * 100)

Data %>%  
  filter(EU > -1 & (sex == "female") & (agegr == "56 to 70")) %>%
  count( sex, EU, agegr) %>% 
  mutate(perc = (n / sum(n)) * 100)







        
```


Write a couple of sentences interpreting your results.

Based on the samples the majority of women voting to remain is much more significant than that of men. 56.73% of women voted to remain while 51.447% of men voted to remain. This shows that when combining all ages of men, the margin between those voting to remain and those voting to leave is fairly small. This can be shown when addressing how people voted based on their age. In both genders the majority of people aged between 16 and 55 voted to remain. The difference comes in ages above that where the majority of men aged 55 to 70 voted to leave while the majority of women voted to stay by a very narrow margin of less than 1% (28 people). The majority of people of both genders aged above 70 voted to leave. These results suggest that the younger the voter is the more likely they will vote to remain.